---
title: "Econometrics Assignement 5 - Brodard Lionel, Giro Larraz Tomas, Marchal Antoine, Schonenberger Ivan"
output: 
  html_notebook: default
---


------->  Please run this code with ' Run all' of *Ctrl+Alt+R*




```{r}
#a)
X_T_X <- matrix(c(100,123,96,109,123,252,125,189,96,125,167,146,109,189,146,168), nrow=4,ncol=4)
y_T_y <- c(3924)
X_T_y <- c(460,810,615,712)
n <- X_T_X[1,1]

cor <- matrix(, nrow=3,ncol=3)

for( i in 1:nrow(cor)){
  for(j in 1:ncol(cor)){
    col_i_mean <- X_T_X[i+1,1]/n
    col_j_mean <- X_T_X[j+1,1]/n

    cor[i,j]<-(X_T_X[i+1,j+1]-n*col_i_mean*col_j_mean)/(sqrt(X_T_X[i+1,i+1]-n*col_i_mean*col_i_mean)*sqrt(X_T_X[j+1,j+1]-n*col_j_mean*col_j_mean))
  }
}

print(cor)
#b)
## high condition number
ei <- eigen(X_T_X)$values

compute_cond<- function(X){
  v <-eigen(X)$values
  sqrt(max(v)/min(v))
}
cond <- compute_cond(X_T_X)

print(cond)


#c)
vifs <-c()
betas <- c()
for(i in 1:3){
  design_k <- solve(X_T_X[-(i+1),-(i+1)])
  val_Xt <- c(X_T_X[i+1,1],X_T_X[i+1,2],X_T_X[i+1,3],X_T_X[i+1,4])
  val_Xt <- val_Xt[-(i+1)]
  beta_k <- design_k %*% val_Xt
  betas <- cbind(betas,beta_k)
  
  above <-0
  below <-0
  if(i==3){
    above <- X_T_X[i+1,i+1]+t(beta_k)%*%X_T_X[-(i+1),-(i+1)]%*%beta_k-2*(beta_k[1]*X_T_X[1,i+1]+beta_k[2]*X_T_X[2,i+1]+beta_k[3]*X_T_X[3,i+1])
    below <- X_T_X[i+1,i+1]-X_T_X[1,i+1]*X_T_X[1,i+1]/n
  }
  else if(i==2){
    above <- X_T_X[i+1,i+1]+t(beta_k)%*%X_T_X[-(i+1),-(i+1)]%*%beta_k-2*(beta_k[1]*X_T_X[1,i+1]+beta_k[2]*X_T_X[2,i+1]+beta_k[3]*X_T_X[4,i+1])
    below <- X_T_X[i+1,i+1]-X_T_X[1,i+1]*X_T_X[1,i+1]/n
  }
  else{
    above <- X_T_X[i+1,i+1]+t(beta_k)%*%X_T_X[-(i+1),-(i+1)]%*%beta_k-2*(beta_k[1]*X_T_X[1,i+1]+beta_k[2]*X_T_X[3,i+1]+beta_k[3]*X_T_X[4,i+1])
    below <- X_T_X[i+1,i+1]-X_T_X[1,i+1]*X_T_X[1,i+1]/n
  }
  r_sq <- 1 - above/below
  vifs <- append(vifs, 1/(1-r_sq))
}

print(vifs)

##vif highest for 3, thus we remove it

##d)

beta1 <- solve(X_T_X)%*%X_T_y
##dropped covariate regression
beta_small <- solve(X_T_X[-4,-4])%*%X_T_y[-4]

print(beta1)
print(beta_small)
```
#Exercice3
#3.a)


#compute the 95% confidence interval for the models parameters
```{r}
df= read.csv('./US_Gasoline_Market.csv')
CONS=1e6*df$GASEXP/df$GASP/df$POP
df$CONS=CONS
#Compute the parameter estimates BetaHat
X=cbind(matrix(replicate(52,1),c(52,1)),df$INCOME,df$GASP,df$PNC,df$PUC,df$PPT,df$PD,df$PN,df$PS,df$YEAR)
Y=matrix(df$CONS,c(52,1))
BetaHat=solve(t(X)%*%X)%*%t(X)%*%Y
#comparing with the linear fit lm
Model=lm(df$CONS~1+df$INCOME+df$GASP+df$PNC+df$PUC+df$PPT+df$PD+df$PN+df$PS+df$YEAR)
summary(Model)
#compute the estimator Ssquared and S matrix
n=nrow(X)
K=ncol(X)
M=diag(n)-X%*%solve(t(X)%*%X)%*%t(X)
E=M%*%Y
Ssquared=t(E)%*%E/(n-K)
S=solve(t(X)%*%X)
#compute the quantile t_(1-alpha),(n-K)
alpha=0.05
t=qt(1-alpha/2,n-K)
#compute the intervals of 95% confidence
Interval=matrix(0L,K,2)
for (k in 1:K){
  Interval[k,1]= BetaHat[k,1]-t*sqrt(Ssquared*S[k,k])
  Interval[k,2]= BetaHat[k,1]+t*sqrt(Ssquared*S[k,k])

}
message("Interval of confidence-Own procedure")
Interval
#verification with confint function, giving the confidence intervals
Int=confint(Model,level=0.95)
message("Interval of confidence-R function")
Int
```

#3.b)

#Compute the correlation between the variables
```{r}
Variables=cbind(df$INCOME,df$GASP,df$PNC,df$PUC,df$PPT,df$PD,df$PN,df$PS,df$YEAR)
cor(Variables)
```
#compute the condition number of the design matrix
```{r}
X_tX=t(X)%*%X
#normalisation
for (k in 1:K ){
 X_tX[,k]=X_tX[,k]/as.vector(sqrt(t(X_tX[,k])%*%X_tX[,k]))
}
Eigenvalues= eigen(X_tX)$values
ConditionNumber=sqrt(max(Eigenvalues)/min(Eigenvalues))
ConditionNumber

```
#Compute vif
```{r}
Rsquared_k= matrix(0L,K-1,1)
 for (k in 1:(K-1)){
   model=lm(Variables[,k]~Variables[,-k])
   Rsquared_k[k,1]=summary(model)$r.squared
 }
message("R_k")
Rsquared_k
VIF=matrix(1L,K-1,1)/(matrix(1L,K-1,1)-Rsquared_k)
message("VIF-Own procedure")
VIF
#Comparison with vif function (package "car")
library(car)
message("VIF-R function")
vif(Model)

```


# Exercice 4
# Consider the dataset and the model of exercise 5 in assignment 4. We are interested in the efficient scale, which is the output Q for which the cost curve reaches its minimum.
# Copy of the exercise 5 in assignment 4.

```{r}
# Importation of Data
df <- read.csv('./Electricity_Generation_Cost.csv')
# Computation of Estimating Equation Variables
df$logCOST_PF<-log(df$COST/df$PF) # log(Total cost/Fuel price)
df$logq <- (log(df$Q)) # log(Total Output)
df$logq2 <- 0.5*(log(df$Q)^2) #1/2* log(Total Output)^0.5
df$logPK_PF<-log(df$PK/df$PF)  #log(Capital price index/Fuel price)
df$logPL_PF<-log(df$PL/df$PF) #log(Wage rate/Fuel price)

# Number of Observation by variable
n<-length(df$logPK_PF)

#### Estimation of parameters, their covariance matrix and their standard errors.
# Matrix of X and Y
X <- cbind(replicate(n,1),df$logq,df$logq2,df$logPK_PF,df$logPL_PF) #We create a matrix with the variable of interest
Y <- cbind(df$logCOST_PF)

# Estimated parameters
Beta_hat <- solve(t(X)%*%X)%*%t(X)%*%Y #parameters estimated with our own procedure

reg <- lm(Y~X[,2:5]) 
Beta_hat_ofs <- cbind(summary(reg)$coef[,1]) #parameters estimated using an off-the-shelve package

# Residuals
e <- Y-X%*%Beta_hat # computation of residuals
ee <- (t(e)%*%e) # sum of squared residuals

# Number of regressors
k<-nrow(Beta_hat) 

# Estimated asymsoptical covariances
covB <-ee[1]/(n-k)*solve(t(X)%*%X) # estimated asymptotical covariance matrix using our own procedure
covB_ofs <- vcov(reg) # estimated asymptotical covariance matrix using an off-the-shelve package

# Estimated standard errors of the parameter estimates
sdB <- cbind(sqrt(diag(covB))) # estimated standard errors using our own procedure
sdB_ofs <- cbind(summary(reg)$coef[,2]) # estimated standard errors using an off-the-shelve package


message('Estimated Coefficient - Own Procedure=')
round(Beta_hat,3)
message('Estimated Coefficient - Off-the-Shelve Package=')
round(Beta_hat_ofs,3)

message('Estimated Covariance Matrix - Own Procedure=')
round(covB,5)
message('Estimated Covariance Matrix - Off-the-Shelve Package=')
round(covB_ofs,5)

message('Estimated Standard Errors - Own Procedure=')
round(sdB,3)
message('Estimated Standard Errors - Off-the-Shelve Package=')
round(sdB_ofs,3)
```

# Exercise 4
# (b) Compute an estimate of Q* using your previous regression results and provide a 95% confidence interval for Q.
```{r}
# We take from our environment the estimates that are useful for us 
Beta <-Beta_hat[2]
Gamma <- Beta_hat[3]
Var_Beta <- covB[2,2]
Sd_Beta <- sqrt(Var_Beta)
Var_Gamma <- covB[3,3]
Sd_Gamma <- sqrt(Var_Gamma)

# Compute the estimation of Q*
Q_star <- exp((1-Beta)/Gamma)

message('Estimated Q*=')
Q_star

######################################################## 
#### Simulation of Q using our estimated parameters ####
######################################################## 

# simulated n
n <-1000
# simulation of 1000 parameters using the estimated mean and sd (be careful of the degree of freedom)
set.seed(10)
Beta_estimation <- rnorm(n,Beta,Sd_Beta/((1/(5-2))*(sqrt(158-2)))) 
set.seed(10)
Gamma_estimation <- rnorm(n,Gamma,Sd_Gamma/((1/(5-2))*(sqrt(158-2))))

# Simulation of Q
Q_estimation <- matrix(1, nrow = n, ncol = n)
for (i in 1:n){
  for (j in 1:n){
  Q_estimation[i,j] <- exp((1-Beta_estimation[i])/Gamma_estimation[j])
  }
}

# Make Q a vector
Q <- as.vector(Q_estimation)

# Find the estimated standard error
message('Standard Error of Q=')
sqrt(var(Q))

# Make a 95% Confidence interval
message('Lower bound =')
Q_star-1.96*sqrt(var(Q))
message('Upper bound=')
Q_star+1.96*sqrt(var(Q))
```
